The Impact of Interview Disruptions on Data Quality in Social Surveys

### 2024-11-26 paper/paper.qmd
update paper and reference

library(arrow)
library(knitr)
library(rstanarm)
library(tidyverse)
library(ggplot2)
library(bayesplot)
library(ggridges)
library(broom.mixed)
library(kableExtra)
library(car)
library(loo)
```

# Introduction

In recent years, the quality of social survey data has become an important topic of inquiry, especially with regard to how different disruptions impact respondent behavior. Several studies have contributed to this body of knowledge by focusing on factors that may affect the reliability of survey responses. Groves et al. (2009) examined the role of interviewer characteristics—such as experience and demeanor—in shaping the quality of responses through a multi-level modeling approach. In another key study, Couper and de Leeuw (2003) investigated how different modes of conducting interviews, like in-person versus telephone, influenced the comfort and attentiveness of respondents. Fowler and Mangione (1990) took a different approach by analyzing observational data to assess the impact of third-party presence during interviews. Although these works laid a valuable foundation for understanding data quality, there remains a gap in systematically analyzing the distinct effects of various kinds of disruptions, such as familial or environmental interference, on respondent understanding and response accuracy.

@@

The remainder of this paper is organized to present a comprehensive exploration of these topics. Section 2 provides a detailed overview of the dataset, focusing on the different variables and their role in assessing data quality. It also outlines the preprocessing steps taken, including data cleaning and the recoding of variables. Section 3 introduces the Bayesian logistic regression methodology, explaining the statistical approach and justifying the selection of priors for the analysis. In Section 4, we present the empirical findings, including parameter estimates and an interpretation of the relationships uncovered by the model. Section 5 discusses the practical implications of these findings, emphasizing the improvements needed in interviewer training and the strategies for reducing respondent interference. Finally, Section 6 summarizes the key contributions of this research, outlines the limitations, and suggests avenues for future work—particularly concerning the integration of new data collection technologies and adaptive survey designs aimed at mitigating the impact of disruptions on respondent data quality.

@@

We use the statistical programming language R (R Core Team 2023) to process, analyze, and visualize survey data exploring the impact of interview disruptions on response quality. The dataset used in this study includes several subsets tailored to different stages of the analysis, enabling a systematic and detailed examination of the research questions. The main dataset for modeling consists of 9,743 rows and 25 variables, capturing detailed information about different types of disruptions (e.g., partner, child, or relative interference), interview methods, and respondent characteristics.

To ensure the reliability of the analysis, we filtered the data to exclude incomplete or ambiguous responses for key variables, such as missing values in "respondent_understood." Responses were also recoded into binary categories to focus on meaningful distinctions in response quality, while variables with excessive categories were grouped to improve interpretability. These preprocessing steps were essential to refine the dataset and facilitate robust modeling.

The following R packages were employed to streamline the data analysis workflow:

-   **tidyverse** (Wickham et al. 2021): Used for efficient data manipulation, cleaning, and transformation. It includes:

-   **dplyr** (Wickham, François, et al. 2021): For summarizing and managing complex datasets.

-   **ggplot2** (Wickham 2021): For creating powerful and flexible visualizations.

-   **rstanarm** (Team 2021): Facilitated the implementation of Bayesian models, enabling straightforward regression modeling using Stan.

-   **arrow** (Apache Arrow 2021): Employed for efficiently reading and writing large datasets, enhancing data handling capabilities across systems.

-   **car** (Fox et al. 2021): Provided functions for regression diagnostics and other statistical analyses, particularly useful for checking assumptions and model validity.

-   **brms** (Bürkner 2021): A package for Bayesian regression modeling, which was used to implement more complex modeling approaches for the data.

-   **rsample** (Kuhn 2021): Used for creating resampling datasets, ensuring robust model validation by facilitating cross-validation and bootstrapping techniques.

-   **janitor** (Firke 2021): Simplified data cleaning processes by helping manage variable names and data structures, making raw data easier to handle.

-   **GGally** (Schloerke et al. 2021): Used to enhance ggplot2 visualizations by providing additional functions for correlation plots and pairwise visualizations, which are useful for exploratory data analysis. This study aims to provide practical insights into how various disruptions affect data quality in survey settings, helping to optimize survey design and execution in the future. \## Measurement

The data for this study is sourced from the European Social Survey (ESS), which collects information on societal attitudes and behaviors across multiple European countries. The dataset focuses on respondents’ understanding of survey questions, their cooperation levels, and the impact of environmental factors such as interruptions during interviews. All responses are collected using standardized methods to ensure comparability across countries.

Different countries may implement localized adaptations in survey administration, which could influence response behaviors. To account for this, country-level metadata is included, allowing for the analysis of cross-national variations.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-aerial_preview
#| tbl-cap: "Survey responses on the impact of interview disruptions across countries in social surveys"

@@

# Select important variables from the model
important_columns <- c(
  "interview_conduct_method", "partner_interference", "child_interference",
  "parent_interference", "relative_interference", "non_relative_interference",
  "interviewer_age", "interviewer_gender", "country", "question_clarification",
  "respondent_reluctant", "respondent_tried_best"
)

@@

# Display the selected data with better column alignment
analysis_data_selected |>
  head(10) |>  # Limit to 10 rows to ensure readability
  kable(booktabs = TRUE, align = "c", col.names = c(
    "Interview Method", "Partner Interference", "Child Interference",
    "Parent Interference", "Relative Interference", "Non-relative Interference",
    "Interviewer Age", "Interviewer Gender", "Country",
    "Question Clarification", "Respondent Reluctant", "Respondent Tried Best"
  )) %>%
  kable_styling(full_width = FALSE)

## Outcome variables

@@

From the plot, we can observe that interference is present across all effort levels, though its proportion varies slightly. Higher levels of effort (e.g., 4 and 5) tend to have a similar distribution of interference presence, while the "Other" category shows a slightly lower proportion of interference.

This visualization provides an overview of how respondent effort and the presence of interference interact in the dataset.

### Key Observations:

-   Interviewers aged 40 to 50 are prominent in levels 4 and 5, showing higher densities in these groups.
-   Clarification levels 1 and 2 show a broader distribution of interviewer ages, with no distinct peak.
-   The "Other" category exhibits a more balanced spread of ages, without significant density concentrations.

This visualization highlights the distribution of interviewer age across clarification levels and provides insights into age-related patterns in survey interactions.

@@

@@

We ran the model in R using the `rstanarm` package (Team, 2021). The default priors from `rstanarm` provide an appropriate level of smoothing and regularization, making the model applicable to survey data with complex patterns while still allowing the data to reveal meaningful relationships, especially concerning factors that affect response quality.

@@

# Results and interpretation

@@

# Discussion

@@

# Model details {#sec-model-details}

## Posterior predictive check
###Validation
```{r}
#| echo: false
#| warning: false

@@

vif(first_model)

residuals <- residuals(first_model, type = "response")
hist(residuals, main = "Residual Histogram", xlab = "Residuals")

r2_values <- bayes_R2(first_model)
r2_df <- data.frame(Index = 1:length(r2_values), R2 = r2_values)

ggplot(r2_df, aes(x = Index, y = R2)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Bayesian R^2 Metrics for first_model Observations", x = "Observation Index", y = "R^2 Value") +
  theme_minimal()

# Perform LOO cross-validation and assign to loo_results
loo_results <- loo(first_model)
# Extract Pareto k values from loo results and plot
pareto_k <- loo_results$diagnostics$pareto_k
pareto_k_df <- data.frame(Index = 1:length(pareto_k), Pareto_k = pareto_k)

loo_df <- data.frame(Observation = 1:length(loo_results$pointwise[,"elpd_loo"]),
                     ELPD_LOO = loo_results$pointwise[,"elpd_loo"])

ggplot(loo_df, aes(x = Observation, y = ELPD_LOO)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(title = "Leave-One-Out Cross Validation (ELPD_LOO)", x = "Observation Index", y = "ELPD_LOO") +
  theme_classic()
```

@@

## Diagnostics

>>>>

library(arrow)
library(knitr)
library(rstanarm)
library(tidyverse)
library(ggplot2)
library(bayesplot)
library(ggridges)
library(broom.mixed)
library(kableExtra)
library(car)
library(loo)
library(modelsummary)
```

# Introduction {#sec-intro}

In recent years, the quality of social survey data has become an important topic of inquiry, especially with regard to how different disruptions impact respondent behavior. Several studies have contributed to this body of knowledge by focusing on factors that may affect the reliability of survey responses. Groves et al. [@groves2011survey] examined the role of interviewer characteristics—such as experience and demeanor—in shaping the quality of responses through a multi-level modeling approach. In another key study, Couper and de Leeuw [@massey2013we] investigated how different modes of conducting interviews, like in-person versus telephone, influenced the comfort and attentiveness of respondents. Fowler and Mangione [@fowler1990standardized] took a different approach by analyzing observational data to assess the impact of third-party presence during interviews. Although these works laid a valuable foundation for understanding data quality, there remains a gap in systematically analyzing the distinct effects of various kinds of disruptions, such as familial or environmental interference, on respondent understanding and response accuracy.

@@

The paper is structured to offer an in-depth examination of the factors that affect respondent comprehension during social survey interviews. Following the introduction in @sec-intro, @sec-data provides a detailed overview of the dataset, focusing on the different variables and their role in assessing data quality. It also outlines the preprocessing steps taken, including data cleaning and the recoding of variables. @sec-model introduces the Bayesian logistic regression methodology, explaining the statistical approach and justifying the selection of priors for the analysis. In @sec-result, we present the empirical findings, including parameter estimates and an interpretation of the relationships uncovered by the model. @sec-discussion discusses the practical implications of these findings, emphasizing the improvements needed in interviewer training and the strategies for reducing respondent interference. Finally, @sec-model-details summarizes the key contributions of this research, outlines the limitations, and suggests avenues for future work—particularly concerning the integration of new data collection technologies and adaptive survey designs aimed at mitigating the impact of disruptions on respondent data quality.

@@

We use the statistical programming language R [@citeR] to process, analyze, and visualize survey data exploring the impact of interview disruptions on response quality. The dataset for this study is sourced from [European Social Survey](https://ess.sikt.no/en/datafile/0c3f5fda-4fe0-42f2-b4c8-4ef093c0508f/7?tab=0&elems=de8c08b6-e406-4c56-92ee-4272818cae96_1), which collects information on societal attitudes and behaviors across multiple European countries. It includes several subsets tailored to different stages of the analysis, allowing a systematic examination of factors affecting survey response quality. The main dataset, used for modeling, consists of 9,743 rows and 25 variables. These data points capture details about various types of disruptions (e.g., partner, child, or relative interference), interview methods, and respondent characteristics. All responses are collected using standardized methods to ensure comparability across countries, thereby facilitating a detailed exploration of the impact of environmental factors, such as interruptions during interviews, on response quality.

To ensure the reliability of the analysis, we filtered the data to exclude incomplete or ambiguous responses for key variables, such as missing values in "respondent_understood." Responses were also recoded into binary categories to focus on meaningful distinctions in response quality, while variables with excessive categories were grouped to improve interpretability.Different countries may implement localized adaptations in survey administration, which could influence response behaviors. To account for this, country-level metadata is included, allowing for the analysis of cross-national variations. These preprocessing steps were essential to refine the dataset and facilitate robust modeling.

The following R packages were utilized to facilitate and streamline the data analysis process:

- **arrow** [@arrow]: Employed for efficiently reading and writing large datasets, enhancing data handling across different systems.

- **knitr** [@knitr]: Used to create dynamic documents with R, including reports and presentations. It enables seamless integration of R code and text.

- **rstanarm** [@rstanarm]: Provides tools for Bayesian regression modeling using the 'Stan' platform, making it straightforward to apply Bayesian models in a regression context.

- **tidyverse** [@tidyverse]: A collection of R packages used for data manipulation, visualization, and cleaning, providing a unified structure to handle complex data workflows.

- **ggplot2** [@ggplot2]: Utilized for creating detailed visualizations, offering a flexible grammar of graphics to develop sophisticated charts and graphs.

- **bayesplot** [@bayesplot]: Facilitates the visualization of Bayesian model results, useful for diagnostic checks and interpreting the results of Bayesian models.

- **ggridges** [@ggridges]: Used to create ridge plots, which are useful for visualizing distributions, enhancing the exploration of posterior distributions from Bayesian models.

- **broom.mixed** [@broom]: Extends the broom package to support models with random effects, summarizing mixed models in a tidy and interpretable format.

- **kableExtra** [@kableExtra]: Enhances the basic table formatting capabilities of the 'knitr' package, providing additional functions for creating aesthetically pleasing and informative tables.

- **car** [@car]: Used for regression diagnostics and statistical analysis, particularly valuable for testing assumptions and evaluating model quality.

- **loo** [@loo]: Provides tools for leave-one-out cross-validation, which was crucial in assessing model performance and ensuring that the Bayesian models generalize well.

- **modelsummary** [@modelsummary]: Summarizes statistical models in a clear, tabular format, supporting the communication of regression and other model results.

- **_Telling Stories with Data_** [@tellingstories]: This book was referenced for its code and methodologies in presenting data and statistical information.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-survey_preview
#| tbl-cap: "Survey interview disruptions and factors influencing respondent comprehension"

@@

# Select important variables from the model
important_columns <- c(
  "respondent_understood_binary", "partner_interference", 
  "parent_interference",  
  "interviewer_age",  "country")

@@

# Display the selected data with better column alignment
analysis_data_selected |>
  head(10) |>  # Limit to 10 rows to ensure readability
  kable(booktabs = TRUE, align = "c", col.names = c(
     "respondent_understood_binary", "partner_interference", 
  "parent_interference",  
  "interviewer_age",  "country"
  )) %>%
  kable_styling(full_width = FALSE)

```
@tbl-survey_preview presents the first six rows from the cleaned dataset, focusing on interview disruptions and respondent comprehension during social survey interviews.
## Outcome variables

@@

From the plot, we can observe that interference is present across all effort levels, though its proportion varies slightly. Higher levels of effort (e.g., 4 and 5) tend to have a similar distribution of interference presence, while the "Other" category shows a slightly lower proportion of interference.This visualization provides an overview of how respondent effort and the presence of interference interact in the dataset.

@@

# Model {#sec-model}

@@

We ran the model in R [@citeR] using the `rstanarm` package@rstanarm. The default priors from `rstanarm` provide an appropriate level of smoothing and regularization, making the model applicable to survey data with complex patterns while still allowing the data to reveal meaningful relationships, especially concerning factors that affect response quality.

@@

# Results {#sec-result}

@@

# Discussion {#sec-discussion}

@@

# Model details {#sec-model-details}
## Model summary
```{r}
#| label: tbl-summary
#| tbl-cap: "Coefficients from a Bayesian logistic regression model examining factors affecting respondents' understanding during interviews, including interview methods, interference types, and country-level indicators."
#| echo: false
#| warning: false
modelsummary(
  list("Bayesian Logistic Regression" = first_model),
  metrics = c("R2", "RMSE")
)
```
@tbl-summary presents the coefficients from our model analyzing factors that influence response comprehension during interviews. The intercept value, 1.823, represents the baseline level of response comprehension when all other predictors are at their reference or baseline levels. Key variables include "interview_conduct_method9," which has a highly positive coefficient of 81.309, indicating that this particular interview method greatly enhances respondent comprehension. In contrast, factors such as "parent_interference" and "child_interference" exhibit negative coefficients (-1.586 and -0.623 respectively), suggesting that these types of interference are associated with lower levels of understanding during interviews.

Additionally, non-relative interference, which shows a positive coefficient of 0.941, suggests that having individuals outside the immediate family present may be less disruptive, or even beneficial, for the comprehension of survey questions. Interviewer-related variables, such as "interviewer_gender," appear to have minimal effects on comprehension, as evidenced by small coefficients like 0.100, suggesting that interviewer attributes play a relatively minor role compared to environmental and relational factors during social surveys
## VIF
```{r}
#| label: tbl-vif
#| tbl-cap: "Variance Inflation Factors (VIF) for Each Predictor in the Model."
#| echo: false
#| warning: false

vif_values <- vif(first_model)

gvif_df <- data.frame(
  Variable = rownames(vif_values), 
  GVIF = round(vif_values[, "GVIF"], 3),
  Df = vif_values[, "Df"],
  `GVIF^(1/(2*Df))` = round(vif_values[, "GVIF^(1/(2*Df))"], 3)
)

kable(gvif_df)

```
@tbl-vif presents the Generalized Variance Inflation Factors (GVIF) for each predictor in the Bayesian logistic regression model. GVIF helps assess multicollinearity among predictors, indicating how much the variance of a regression coefficient is inflated due to correlations between variables.

Most predictors, such as respondent_tried_best (GVIF = 3.186, Df = 5) and respondent_reluctant (GVIF = 2.725, Df = 5), show moderate GVIF values, suggesting manageable multicollinearity. The adjusted GVIF values, which are generally close to 1, indicate that the inflation effect is controlled across variables.

In summary, the GVIF analysis indicates that multicollinearity is present but not severe enough to threaten model reliability. Further analysis may focus on predictors with slightly higher GVIF to ensure stability.

## Posterior predictive check
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior Predictive Check: The comparison between observed response values (dark line) and model-generated replications (light lines) demonstrates the model's ability to replicate the observed response patterns, indicating a good fit and capturing the main features of the response distribution."

@@

```
In @fig-ppcheck, we implement a posterior predictive check, which displays the overlap between the observed data (denoted by \( y \)) and the replicated data generated from the model (denoted by \( y_{\text{rep}} \)). The replicated lines closely follow the shape and central trend of the observed data, particularly around the main peak at 0.9, which suggests that the model captures the overall distribution effectively. This alignment indicates that the model successfully represents the underlying data-generating process, especially in regions with high data density. While the model's performance appears strong, any visible discrepancies—particularly in areas of lower density—may highlight aspects of the data structure that are not fully captured. Addressing such discrepancies could involve refining the model by introducing additional parameters or reconsidering underlying assumptions. Overall, this comparison provides evidence that the model fits well, though there remain potential areas for enhancement to better encompass the full range of data complexities.

@@

## Diagnostics
### Validation
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-diagnostic-checks
#| fig-cap: "Comparing Diagnostic Visualizations for Model Checking"
#| fig-subcap: ["Residual Histogram", "Pareto k Diagnostic Values"]
#| layout-ncol: 2

# Generate Residual Histogram
residuals <- residuals(first_model, type = "response")
hist(residuals, main = "Residual Histogram", xlab = "Residuals")

# Perform LOO cross-validation and assign to loo_results
loo_results <- loo(first_model)

# Extract Pareto k values from loo results and create the Pareto k Diagnostic plot
pareto_k <- loo_results$diagnostics$pareto_k
pareto_k_df <- data.frame(Index = 1:length(pareto_k), Pareto_k = pareto_k)

```
In Figure @fig-diagnostic-checks-1, the residual histogram displays the distribution of residuals from the model. The concentrated frequency around zero suggests that the model effectively captures much of the variability in the data, with fewer extreme deviations, thereby indicating a reasonable level of fit.

Figure @fig-diagnostic-checks-2 presents the Pareto k diagnostic values, which help assess the reliability of the estimated leave-one-out cross-validation (LOO) predictions. The majority of Pareto k values fall below the threshold of 0.7, indicating that the model is well-calibrated with only minimal issues in terms of influential observations.
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-loo-rhat-checks
#| fig-cap: "Comparing Diagnostic Visualizations for Model Checking"
#| fig-subcap: ["Leave-One-Out Cross Validation (ELPD_LOO)", "Rhat Diagnostic Plot"]
#| layout-ncol: 2
# Create Leave-One-Out Cross Validation (ELPD_LOO) plot
loo_df <- data.frame(Observation = 1:length(loo_results$pointwise[,"elpd_loo"]),
                     ELPD_LOO = loo_results$pointwise[,"elpd_loo"])

ggplot(loo_df, aes(x = Observation, y = ELPD_LOO)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(title = "Leave-One-Out Cross Validation (ELPD_LOO)", x = "Observation Index", y = "ELPD_LOO") +
  theme_classic()

# Generate Rhat Diagnostic Plot
rhat_values <- rhat(first_model)
mcmc_rhat(rhat_values)

Figure @fig-loo-rhat-checks-1 is a scatter plot for leave-one-out cross-validation (ELPD_LOO) estimates. The consistency of the points around the red loess smooth line suggests that the model’s predictions are largely consistent across different observations, thus reinforcing the model's robustness.

Figure @fig-loo-rhat-checks-2 is an Rhat plot, showing that all values of the Rhat diagnostic are close to 1. This indicates good mixing of the Markov Chain Monte Carlo (MCMC) chains and suggests that the model has likely converged appropriately, making the posterior estimates trustworthy.
